---
title: "Data Science Capstone"
subtitle: "Sentence Fragments and NLP I"
author: "Rohan Lewis"
date: "2020.12.09"
output:
  html_document :
    fig_height : 5
    fig_width : 8
    highlight : espresso
  
---

* * *

# I.  Background

I am now using the lines from all three documents to predict the next word for a quiz.

This data will be eventually used for predicting the next word in a random statement.

See the Appendix for code.

* * *

```{r echo = FALSE, message = FALSE}

knitr::opts_chunk$set(comment = NA)
options("scipen" = 100)
options(java.parameters = "-Xmx8g")

```

```{r echo = FALSE, message = FALSE}

library(dplyr)
library(readr)
library(RWeka)
library(stringr)
library(textclean)
library(tm)

```

```{r read, echo = FALSE, message = FALSE}
#Files are read.

blogs <- read_lines(file = "en_US.blogs.txt")

news <- read_lines(file = "en_US.news.txt")

tweets <- read_lines(file = "en_US.twitter.txt")

```

```{r functions, echo = FALSE, message = FALSE}

#Function:
#"’" is replaced with "'", contractions are expanded.
#"U.S." is replaced with "US".
#"a.m." is replaced with "am".
#"p.m." is replaced with "pm".
replacement <- function(data) {
  
  temp <- replace_contraction(gsub("’", "'", data),
                              contraction.key = lexicon::key_contractions)
  temp <- gsub("U.S.", "US", temp)
  temp <- gsub("a.m.", "am", temp)
  temp <- gsub("p.m.", "pm", temp)
  return(temp)
}



#Function returns the word immediately after a string for all lines that contain the string."
retrieve_next_words <- function(input_string) {

  #Word length of input string.
  n <- length(strsplit(input_string, " ")[[1]])
  #Final list.
  next_words <- NULL
  
  for (bnt in bnts) {
    #Check if the string is in the line.
    if(grepl(input_string, bnt)) {
      #Find where in the line it is located.
      beginning <- str_locate(bnt, input_string)[1]
      #Start from where the string is, split by space, grab the next word.
      next_word <- strsplit(substring(bnt, beginning), " ")[[1]][n+1]
      #Get rid of non-alphanumeric.
      next_word <- str_replace_all(next_word, regex("\\W+"), "")
      #Update final list.
      next_words <- c(next_words, next_word)
    }
  }
  return(next_words)
}



possibilities <- function(q_string){
  
  #Cleanup.
  q_string <- replacement(q_string)
  #Word length of input string.
  temp_str <- str_split(q_string, " ")[[1]]
  n <- length(temp_str)
  
  #Start with last 10 words.
  if (n > 10) {
    q_string <- paste(temp_str[(n-9):n], collapse = ' ')    
  }

  while (nchar(q_string > 0)) {

    #Get next words from blogs, news, and tweets.
    q_words <- retrieve_next_words(q_string)
    
    #If there is at least one word...
    if (length(q_words > 0)) {
      
      #Return at most 10 entries of the frequency table.
      df_freq <- df_pred_creator(q_words)
      
      if (dim(df_freq)[1] > 10){
        return(df_freq[1:10,])
      } else {
        return(df_freq)
      }
      
    #No words means try again, with a shortened string.  
    } else {
      #Shorten string by removing left most word.
      q_string <- substring(q_string, str_locate(q_string, " ")[1]+1)
    }
  }
  #All string lengths failed to match.
  return("NO MATCHES")
}



#This function will create a Frequency dataframe from next_words.
df_pred_creator <- function(next_words) {

  #Convert sorted table to dataframe.
  df <- as.data.frame(sort(table(next_words), decreasing = T))
  
  #If there is only one row, change to two columns from index and single column.
  if (length(next_words) == 1) {
    df <- cbind(newColName = rownames(df), df)
    rownames(df) <- 1:nrow(df)
  }
  
  #Change Column Names.
  df <- df %>%
    `colnames<-`(c("Possible Next Word", "Frequency"))
  
  return(df)
}

print("SETUP COMPLETE")
```

* * *

# II. Predictions

```{r predictions, echo = FALSE, message = FALSE}

blogs <- replacement(blogs)

news <- replacement(news)

tweets <- replacement(tweets)

#List of all blogs, news, and tweets.
bnts <- c(blogs, news, tweets)

print("REPLACEMENT COMPLETE")

```

## 1. "The guy in front of me just bought a pound of bacon, a bouquet, and a case of..."

```{r q1, echo = FALSE, message = FALSE}

q1_string <- "The guy in front of me just bought a pound of bacon, a bouquet, and a case of"

possibilities(q1_string)

```

## 2. "You're the reason why I smile everyday. Can you follow me please? It would mean the..."

```{r q2, echo = FALSE, message = FALSE}

q2_string <- "You're the reason why I smile everyday. Can you follow me please? It would mean the"

possibilities(q2_string)

```

## 3. "Hey sunshine, can you follow me and make me the..."

```{r q3, echo = FALSE, message = FALSE}

q3_string <- "Hey sunshine, can you follow me and make me the"

possibilities(q3_string)

```

## 4. "Very early observations on the Bills game: Offense still struggling but the..."

```{r q4, echo = FALSE, message = FALSE}

q4_string <- "Very early observations on the Bills game: Offense still struggling but the"

possibilities(q4_string)

```

## 5. "Go on a romantic date at the..."

```{r q5, echo = FALSE, message = FALSE}

q5_string <- "Go on a romantic date at the"

possibilities(q5_string)

```

## 6. "Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my..."

```{r q6, echo = FALSE, message = FALSE}

q6_string <- "Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my"

possibilities(q6_string)

```

## 7. "Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some..."

```{r q7, echo = FALSE, message = FALSE}

q7_string <- "Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some"

possibilities(q7_string)

```

## 8. "After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little..."

```{r q8, echo = FALSE, message = FALSE}

q8_string <- "After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little"

possibilities(q8_string)

```

## 9. "Be grateful for the good times and keep the faith during the..."

```{r q9, echo = FALSE, message = FALSE}

q9_string <- "Be grateful for the good times and keep the faith during the"

possibilities(q9_string)

```

## 10. "If this isn't the cutest thing you've ever seen, then you must be..."

```{r q10, echo = FALSE, message = FALSE}

q10_string <- "If this isn't the cutest thing you've ever seen, then you must be"

possibilities(q10_string)

```

# Appendix

* * *

## Setup

```{r eval = FALSE}

knitr::opts_chunk$set(comment = NA)
options("scipen" = 100)
options(java.parameters = "-Xmx4g")

```

### Packages

```{r eval = FALSE}

library(dplyr)
library(readr)
library(RWeka)
library(stringr)
library(textclean)
library(tm)

```

### Read Files

```{r eval = FALSE}
#Files are read.

blogs <- read_lines(file = "en_US.blogs.txt")

news <- read_lines(file = "en_US.news.txt")

tweets <- read_lines(file = "en_US.twitter.txt")
```

### Helper Functions

```{r eval = FALSE}

#Function:
#"’" is replaced with "'", contractions are expanded.
#"U.S." is replaced with "US".
#"a.m." is replaced with "am".
#"p.m." is replaced with "pm".
replacement <- function(data) {
  
  temp <- replace_contraction(gsub("’", "'", data),
                              contraction.key = lexicon::key_contractions)
  temp <- gsub("U.S.", "US", temp)
  temp <- gsub("a.m.", "am", temp)
  temp <- gsub("p.m.", "pm", temp)
  return(temp)
}



#Function returns the word immediately after a string for all lines that contain the string."
retrieve_next_words <- function(input_string) {

  #Word length of input string.
  n <- length(strsplit(input_string, " ")[[1]])
  #Final list.
  next_words <- NULL
  
  for (bnt in bnts) {
    #Check if the string is in the line.
    if(grepl(input_string, bnt)) {
      #Find where in the line it is located.
      beginning <- str_locate(bnt, input_string)[1]
      #Start from where the string is, split by space, grab the next word.
      next_word <- strsplit(substring(bnt, beginning), " ")[[1]][n+1]
      #Get rid of non-alphanumeric.
      next_word <- str_replace_all(next_word, regex("\\W+"), "")
      #Update final list.
      next_words <- c(next_words, next_word)
    }
  }
  return(next_words)
}



possibilities <- function(q_string){
  
  #Cleanup.
  q_string <- replacement(q_string)
  #Word length of input string.
  temp_str <- str_split(q_string, " ")[[1]]
  n <- length(temp_str)
  
  #Start with last 10 words.
  if (n > 10) {
    q_string <- paste(temp_str[(n-9):n], collapse = ' ')    
  }

  while (nchar(q_string > 0)) {

    #Get next words from blogs, news, and tweets.
    q_words <- retrieve_next_words(q_string)
    
    #If there is at least one word...
    if (length(q_words > 0)) {
      
      #Return at most 10 entries of the frequency table.
      df_freq <- df_pred_creator(q_words)
      
      if (dim(df_freq)[1] > 10){
        return(df_freq[1:10,])
      } else {
        return(df_freq)
      }
      
    #No words means try again, with a shortened string.  
    } else {
      #Shorten string by removing left most word.
      q_string <- substring(q_string, str_locate(q_string, " ")[1]+1)
    }
  }
  #All string lengths failed to match.
  return("NO MATCHES")
}



#This function will create a Frequency dataframe from next_words.
df_pred_creator <- function(next_words) {

  #Convert sorted table to dataframe.
  df <- as.data.frame(sort(table(next_words), decreasing = T))
  
  #If there is only one row, change to two columns from index and single column.
  if (length(next_words) == 1) {
    df <- cbind(newColName = rownames(df), df)
    rownames(df) <- 1:nrow(df)
  }
  
  #Change Column Names.
  df <- df %>%
    `colnames<-`(c("Possible Next Word", "Frequency"))
  
  return(df)
}

print("SETUP COMPLETE")
```

# II. Predictions

```{r eval =  FALSE}

blogs <- replacement(blogs)

news <- replacement(news)

tweets <- replacement(tweets)

bnts <- c(blogs, news, tweets)

print("REPLACEMENT COMPLETE")

```

### 1.

```{r eval = FALSE}

q1_string <- "The guy in front of me just bought a pound of bacon, a bouquet, and a case of"

possibilities(q1_string)

```

### 2.

```{r eval = FALSE}

q2_string <- "You're the reason why I smile everyday. Can you follow me please? It would mean the"

possibilities(q2_string)

```

### 3.

```{r eval = FALSE}

q3_string <- "Hey sunshine, can you follow me and make me the"

possibilities(q3_string)

```

### 4.

```{r eval = FALSE}

q4_string <- "Very early observations on the Bills game: Offense still struggling but the"

possibilities(q4_string)

```

### 5.

```{r eval = FALSE}

q5_string <- "Go on a romantic date at the"

possibilities(q5_string)

```

### 6.

```{r eval = FALSE}

q6_string <- "Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my"

possibilities(q6_string)

```

### 7. 

```{r eval = FALSE}

q7_string <- "Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some"

possibilities(q7_string)

```

### 8.

```{r eval = FALSE}

q8_string <- "After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little"

possibilities(q8_string)

```

### 9.

```{r eval = FALSE}

q9_string <- "Be grateful for the good times and keep the faith during the"

possibilities(q9_string)

```

### 10.

```{r eval = FALSE}

q10_string <- "If this isn't the cutest thing you've ever seen, then you must be"

possibilities(q10_string)

```